---
title: "Patrick Schöngrundner: GLMM Simulations"
subtitle: "Using MixedModels.jl"
author: "Reinhold Kliegl"
date: "2022-09-09 (last revised: `r format(Sys.time())`)"
format:
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    number-sections: true
    fig-width: 8
    fig-height: 6
editor_options: 
  chunk_output_type: console
---

# Preprocessing

## Packages

```{r}
library(easystats)
library(tidyverse)
library(lme4)
```

## Read simulated data

```{r}
dat1 <- 
  read_delim("./data/DataEx1.csv", col_types = "cfcfliii") |> 
  rename(Subj = subj_ID, Sex = subj_sex, Item = item_ID, Cat = item_cat, GM=`(Intercept)`) |> 
   mutate(
    Subj = factor(paste0("S", str_pad(Subj, width = 2, side = "left", pad = "0"))),
    Item = factor(paste0("I", str_pad(Item, width = 2, side = "left", pad = "0"))),
    ) |> 
  as_tibble()
```

# Contrasts

```{r}
# contrasts for accuracy data
contrasts(dat1$Sex) <- -contr.sum(2)  
contrasts(dat1$Cat) <- contr.sum(2)  
```

# GLMMs

## Factor contrast

```{r}
fit1a <- glmer(resp ~ 1 + Sex + Cat + Sex:Cat + (1 | Subj) + (1 | Item), 
                    family = binomial, data = dat1,
                    control = glmerControl(calc.derivs=FALSE))
print(summary(fit1a), corr=FALSE)
```

Dropping the `Sex` main effect leads to estimating the effexct of `Sex` within the levels of the factor `Cat`, that is a nested re-parameterization with the same number of degrees of freedom. 

```{r}
fit1b <- glmer(resp ~ 1 + Cat + Sex:Cat + (1 | Subj) + (1 | Item), 
                    family = binomial, data = dat1,
                    control = glmerControl(calc.derivs=FALSE))
print(summary(fit1b), corr=FALSE)
```

We can see this also in the LRT. 

```{r}
anova(fit1b, fit1a)
```

The model reparameterization does not change the goodness of fit. 

## Source: PS 

```{r}
fit2a <- glmer(resp ~ 1 + sex1 + cat1 + sex1:cat1 + (1 | Subj) + (1 | Item), 
                    family = binomial, data = dat1,
                    control = glmerControl(calc.derivs=FALSE))
print(summary(fit2a), corr=FALSE)
```

Without factors there are no levels. The model estimates the two coefficients we ask for.

```{r}
fit2b <- glmer(resp ~ 1        + cat1 + sex1:cat1 + (1 | Subj) + (1 | Item), 
                    family = binomial, data = dat1,
                    control = glmerControl(calc.derivs=FALSE))
print(summary(fit2b), corr=FALSE)
```

Now the model degrees of freedom are different and we also have a meaningful LRT

```{r}
anova(fit2b, fit2a)
```

## Source: Model matrix  

We can get indicators also from the model matrix.

```{r}
mm1 <- model.matrix(~ 1 + Sex*Cat, data=dat1)
dat1$sex <- mm1[, 2]
dat1$cat <- mm1[, 3]

fit3a <- glmer(resp ~ 1 + sex + cat + sex:cat + (1 | Subj) + (1 | Item), 
                    family = binomial, data = dat1,
                    control = glmerControl(calc.derivs=FALSE))
print(summary(fit3a), corr=FALSE)
```


```{r}

fit3b <- glmer(resp ~ 1       + cat + sex:cat + (1 | Subj) + (1 | Item), 
                    family = binomial, data = dat1,
                    control = glmerControl(calc.derivs=FALSE))
print(summary(fit3b), corr=FALSE)
```

```{r}
anova(fit3b, fit3a)
```

# Appendix

```{r}
sessionInfo()
```

