---
title: "Pan et al. (2013). Eye–voice span during rapid automatized naming of digits and ￼dice in Chinese normal and dyslexic children"
subtitle: "RePsychLing in SMLP2022"
author: "Reinhold Kliegl"
date: "2022-09-15 (last revised: `r format(Sys.time())`)"
format: 
  html:
    embed-resources: true
    toc: true
    toc-depth: 2
    code-fold: false
    number-sections: true
    fig-width: 8
    fig-height: 6
    fig-format: svg
editor_options: 
  chunk_output_type: console
---

# Background

Predicting psychometric digit-RAN and dice-RAN with gaze duration and 
eye-voice span (EVS) from computerized assessment of digit-RAN and dice-RAN 
in Chinese control and dyslexic children. These results are reported in:

Pan et al. (2013). Eye–voice span during rapid automatized naming of digits  
and ￼dice in Chinese normal and dyslexic children. `Developmental Science`. 

# Setup

## Packages

```{r}
library(arrow)
library(easystats)
library(tidyverse)
library(lme4)
library(latticeExtra)
library(RePsychLing)
library(remef)
library(grid)

vplayout <- function(x, y) {
viewport(layout.pos.row = x, layout.pos.col = y) 
}

theme_set(theme_bw())
# respecting color vision deficiency
cbPalette <- c( "#0072B2", "#D55E00", "#009E73", "#CC79A7",
                "#F0E442", "#56B4E9", "#999999", "#E69F00")
```

## Reading data and preprocessing

```{r}
load("./data/Pan_RAN.rda")

dat <- 
  dat |> 
  tibble() |> 
  select("Subj"="subj", "Group"="group", "Condition"="condition", ran, gaze, evs) |>
  mutate(
    Subj = factor(paste0("S", str_pad(Subj, width = 3, side = "left", pad = "0"))),
    Group = factor(Group, labels=c("control", "dyslexic")),
    Condition = factor(Condition, labels=c("digit", "dice")),
    lran = log(ran),
    gaze_c = gaze - 406,
    evs_c  = evs - .84
  )
```

# LMMs

## Contrasts and indicators

```{r}
contrasts(dat$Group) <- contr.sum(2)
contrasts(dat$Condition) <- contr.sum(2)

mm <- model.matrix(~Group*Condition,data=dat)
dat$grp <- mm[,2]
dat$cnd <- mm[,3]
```


```{r}
m_lme4 <- lmer(ran ~ 1 + (Condition+Group+evs_c+gaze_c)^2 +
                          Group:Condition:evs_c + Group:Condition:gaze_c +
                          (1|Subj), data=dat, REML=FALSE,
                          control=lmerControl(calc.derivs=FALSE))
print(summary(m_lme4), corr=FALSE)
```

# Unadjusted and partial plots of Ran

##  Figure 1: Eye-voice span x Condition x Group
 
```{r}
# Zero-order relation
p1a <- 
  dat |> 
  ggplot(aes(y=ran, x=evs,  group=Condition:Group, color=Condition:Group)) + 
  geom_point() + geom_smooth(method="lm") +
  xlab("Eye-voice span") + 
  scale_y_continuous("RAN", limits=c(0, 35)) + 
  scale_color_manual("Condition x Group", values=cbPalette) +
  theme(legend.position = "none", 
        panel.background=element_rect(fill = "white"))

# Partial effects
dat$CGe1 <- keepef(m_lme4, fix = c(2:4,6:7, 9, 12), ran = NULL) 

p1b <- 
  dat |> 
  ggplot(aes(y=CGe1, x=evs, group=Condition:Group, color=Condition:Group)) +
  geom_point() + geom_smooth(method="lm") +
  xlab("Eye-voice span") + 
  scale_y_continuous("Adjusted RAN", limits=c(0,35)) + 
  scale_color_manual("Condition x Group", values=cbPalette) +
  theme(legend.position = c(.01, .01), legend.justification = c(.01, .01), 
        legend.text = element_text(size=8), 
        panel.background=element_rect(fill = "white")) 

#  -- evs only predictor for control in digit-RAN!

grid.newpage() # Figure 1
pushViewport(viewport(layout = grid.layout(1,2))) 
print(p1a, vp=vplayout(1,1))
print(p1b, vp=vplayout(1,2))
```

# Transfer to Julia

```{r}
write_feather(dat, "./data/Pan_RAN.arrow")
```

# Appendix

```{r}
sessionInfo()
```


