---
title: "Two attributes of numerical meaning"
author: "Oliver Lindemann"
output:
  html_document:
    toc: true
---

```{r libs, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
```

# Description

The study aims to examine the interactions of space- and size-congruity effects while number processing.

The numbers 1 to 9 ($N$) were presented in different six font sizes ($V_\text{size}=[-3, -2, -1, 1, 2, 3]$) and at six different stimulus locations (three left and three right from screen center, $V_\text{space}=[-3, -2, -1, 1, 2, 3]$). For each trial, a size and space-congruity parameter, $C$, could be calculated that varies between $-1$ and $+1$ (see formula below).

The mixed effect model of the reaction times aims to consider, beside these two congruity parameters and their interaction, also the size- or space-congruity in the previous trial ($n-1$, 'Gratton effect').

Congruity was defined as $$C =  \frac{V}{3}  \cdot \frac{N-5}{4} = \frac{V(N-5)}{12}$$

```{r def congruity, echo=FALSE}
congruity_parameter = function(num_size, variable) {
  n = num_size
  v = variable
  return(v*(n-5)/12)
}
```

# Data

## Code book (of the relevant vaiables)

-   **`Subject`**: Subject ID
-   **`digit`**: Presented digit (1 to 9, except 5)
-   **`rt`**: Reaction time in ms
-   **`size_rc`**: Stimulus size (-3 to +3) (recalculated)
-   **`pos_rc`**: Stimulus position (-3 to +3) (recalculated)
-   **`numerical_distance`**: Distance to 5 (abs(5-digit))
-   **`SiC`**: Size congruity parameter (-1 to +1)
-   **`SpC`**: Spatial congruity parameter (-1 to +1)
-   **`P_SiC`**: Size congruity of the previous trial (-1 to +1)
-   **`P_SpC`**: Spatial congruity of the previous trial (-1 to +1)

```{r, echo=FALSE}
# Data proprocessing

raw = read_csv("./data/WBL16_raw_data.csv", na =c("NA", "None"), show_col_types = FALSE) %>%
    mutate(Subject = as_factor(Subject),
           resp = factor(resp, levels=c(32, 105, 111), labels=c("no", "ti", "to")), 
           mapping = as_factor(mapping)
           )
# filter and sort and create new variables
data = raw %>%
    filter(trial>=0) %>%
    arrange(Subject, trial) %>%
    mutate(
      magnitude = factor(digit >5, levels=c(FALSE,TRUE),
                        labels=c("small", "large")),
      pos_rc = pos/100, # pos recoded
      size_rc = as.integer(as.factor(size))-4, # size recoded to -3,-2,-1,1,2,3
      size_rc = ifelse(size_rc>=0, size_rc+1, size_rc),
      numerical_distance = abs(digit-5),
      SiC = congruity_parameter(digit, size_rc),
      SpC = congruity_parameter(digit, pos_rc)
      )

# Determine congruity of previous trails (Gratton effect)
# previous size and space congruence
data = data %>% 
  mutate(P_SiC = append(NA, SiC[2:n()-1]),
         P_SpC = append(NA, SpC[2:n()-1]))

# set first "previous congruity" of each subject to NaN
for (x in unique(data$Subject)) {
  t = min(subset(data, Subject==x)$trial)
  idx = which((data$Subject==x) & (data$trial==t))
  data[idx,"P_SiC"]  = NA
  data[idx,"P_SpC"]  = NA
}


## Select data
rt.data = data %>%
  filter(error==0, resp !="no", rt>200) %>%
  select(-trial, -resp, -magnitude, -mapping, -error, -ISI)
```

## Show data

```{r}
glimpse(rt.data)
```

# Linear Mixed Effects Model Analyses

## Model Selection

```{r eval=TRUE}
# initial model with control factors and factors of interest
model.no_slopes = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                         (1 | Subject),  
                         REML =FALSE, data =rt.data)

# testing random slope for numerical distance
model.nd = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.no_slopes, model.nd)

# testing random slope for visual size
model.nd_vs = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance + size_rc | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.nd, model.nd_vs)

# testing random slope for position
model.nd_loc = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance + pos_rc | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.nd, model.nd_loc)

# testing random slope for SiC
model.nd_sic = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance + SiC | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.nd, model.nd_sic)

# testing random slope for spatial congruity
model.nd_spc = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance + SpC | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.nd, model.nd_spc)

# testing random slope for previous size congruity
model.nd_psic = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance + P_SiC | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.nd, model.nd_psic)

# testing random slope for previous position congruity
model.nd_pspc = lmer(rt~ numerical_distance + size_rc + pos_rc +
                         SiC*SpC*P_SiC*P_SpC +
                         SiC:SpC +
                         SiC:P_SiC +
                         SiC:P_SpC +
                         SpC:P_SiC +
                         SpC:P_SpC +
                        (1 + numerical_distance + P_SpC | Subject),  
                       REML =FALSE, data =rt.data)
anova(model.nd, model.nd_pspc)
```

## Result of best fitting model

```{r best_model}
#### Best model with random slopes for numerical distance only 
summary(model.nd)
```

# Plotting

```{r, echo=FALSE}
rt.data = rt.data %>%
    mutate(
      pos_cat = factor(pos > 0, levels=c(FALSE, TRUE),
                        labels=c("left", "right")),
      size_cat = factor(size > 100, levels=c(FALSE, TRUE),
                        labels=c("small", "large"))
    )


m = length(unique(rt.data$digit)) * length(unique(rt.data$size_cat)) *
              length(unique(rt.data$pos_cat))
tmp = rt.data %>%
          group_by(Subject, digit, size_cat, pos_cat) %>%
          summarise(mRT= median(rt)) %>%
          # norm RT dfor within CI
          group_by(Subject) %>%
          mutate(mRT.norm = mRT - mean(mRT)) %>%
          ungroup() %>%
          mutate(mRT.norm = mRT.norm + mean(mRT)) %>% ## add grand mean
          # averge over subjects and calc CI  
          group_by(digit, size_cat, pos_cat) %>%
          summarize(mean_rt = mean(mRT),
                    n = length(mRT),
                    sd=sd(mRT),
                    se=sd/sqrt(n),
                    ci.between = se*qt(.975, n-1),
                    ci.within = sd(mRT.norm)/sqrt(n)*qt(.975, n-1) * sqrt(m/(m-1)))  

tmp = tmp %>%
    mutate(Condition = paste0(as.character(pos_cat), as.character(size_cat)))

tmp$Condition = str_replace(tmp$Condition, "leftsmall", "Left, Small")
tmp$Condition = str_replace(tmp$Condition, "leftlarge", "Left, Large")
tmp$Condition = str_replace(tmp$Condition, "rightsmall", "Right, Small")
tmp$Condition = str_replace(tmp$Condition, "rightlarge", "Right, Large")

dodge <- position_dodge(.5)
ggplot(data=tmp, aes(x=digit, y=mean_rt, shape=Condition))  +
  geom_line(position=dodge, aes(linetype = Condition)) +
  geom_errorbar(position=dodge, aes(ymin=mean_rt- ci.within , ymax=mean_rt+ ci.within ),
                  width=.8) +
  geom_point(position=dodge, size = 4, aes(shape=Condition, fill=Condition)) +
  scale_fill_manual(values = c("black", "white","black", "white")) +   
  scale_shape_manual(values = c(24, 24, 21,21)) +   
  scale_linetype_manual(values = c(3,3,3,3)) +   
  scale_x_continuous(breaks = c(1, 2,3,4,6,7,8, 9) ) +
  ylab("Mean RT (ms)") +
  xlab("Digit") +
  theme_bw(base_size=18)
```
