---
title: "Exp3 R Markdown"
author: "Özge Bakay"
date: "2022-01-26"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(magrittr)
library(dplyr)
library(tidyr)
library(ggplot2)
#library(rethinking)
library(lme4)
library(lmerTest)
library(ordinal)
library(readxl)
library(sciplot)

df <- read_excel("./data/Bakay_Exp3 R Data.xlsx")
```

## About the data

The data here is from a rating experiment. The experiment investigates whether there is missing-verb illusion in Turkish center-embeddings and whether case similarity has an effect on the presence of illusion. The experiment has a 2x2 design where the variables are Grammaticality (Grammatical vs. Ungrammatical) and Case Similarity (High vs. Low). 

## Accuracy Check

```{r}
df$AccuracyCOMP %<>% as.integer()

accuracy_list <- 
  df %>% 
  dplyr::select(AccuracyCOMP, Subject) %>% 
  group_by(Subject) %>% 
  summarise(Accuracy = mean(AccuracyCOMP))

View(accuracy_list) ## Subject 53 below 70%

## creating a new df by excluding the subject with low accuracy
df_new <- subset(df, Subject != "53")
str(df_new)
unique(df_new$Subject)

## accuracy check again
accuracy_list2 <- 
  df_new %>% 
  dplyr::select(AccuracyCOMP, Subject) %>% 
  group_by(Subject) %>% 
  summarise(Accuracy = mean(AccuracyCOMP))

View(accuracy_list2)

mean(accuracy_list2$Accuracy) 
## 89% overall mean accuracy in the remaining 56 participants

```

## DESCRIPTIVES for critical items

```{r}
df_critical <- df_new %>% subset(Type == "exp")
unique(df_critical$Grammaticality)
unique(df_critical$Case)

## mean of ratings
mean <- 
  df_critical %>% 
  group_by(Case, Grammaticality) %>% 
  dplyr::summarize(RatingRATING = mean(RatingRATING, na.rm =T))

View(mean)

## se of ratings, does not work now for some reason
std <- df_critical %>% group_by(Case, Grammaticality) %>% 
  dplyr::summarize(RatingRATING = se(RatingRATING, na.rm =T))

View(std) 

## comprehension per condition
meanC <- df_critical %>% group_by(Case, Grammaticality) %>% 
  dplyr::summarize(Accuracy = mean(AccuracyCOMP, na.rm =T))

View(meanC)

```

## sum contrast coding

```{r}
df_critical %<>% mutate(Gra_contrast = ifelse(Grammaticality == "Gra", 0.5, -0.5))
df_critical %<>% mutate(Case_contrast = ifelse(Case == "Low", 0.5, -0.5))

```

## clm models with random intercepts only

```{r}
exp4 <- clmm(as.factor(RatingRATING) ~ Trial + Case_contrast * Gra_contrast + (1|Subject) + (1|Item), data=df_critical)
summary(exp4) # sig effect of gra

## split

group = split(df_critical, df_critical$Case_contrast)
dfHigh = group$"-0.5"
dfLow = group$"0.5"

expHigh <- clmm(as.factor(RatingRATING) ~ Trial + Gra_contrast + (1|Subject) + (1|Item), data=dfHigh)
summary(expHigh)#sig

expLow <- clmm(as.factor(RatingRATING) ~ Trial + Gra_contrast + (1|Subject) + (1|Item), data=dfLow)
summary(expLow) #sig, more sig

```

## clm models including random slopes

```{r}
expc <- clmm(as.factor(RatingRATING) ~ Trial + Case_contrast * Gra_contrast + (1+ Case_contrast * Gra_contrast|Subject)+(1+ Case_contrast * Gra_contrast|Item), data=df_critical)
summary(expc) # this model converges but the coefficients are NA. I do not know if we can trust this model.

## split

group = split(df_critical, df_critical$Case_contrast)
dfHigh = group$"-0.5"
dfLow = group$"0.5"

expHigh2 <- clmm(as.factor(RatingRATING) ~ Trial + Gra_contrast + (1+ Gra_contrast |Subject)+(1+ Gra_contrast |Item), data=dfHigh)
summary(expHigh2) #sig effect of gra

expLow2 <- clmm(as.factor(RatingRATING) ~ Trial + Gra_contrast + (1+ Gra_contrast |Subject)+(1+ Gra_contrast |Item), data=dfLow)
summary(expLow2) #sig effect of gra, more sig

group = split(df_critical, df_critical$Gra_contrast)
dfGra = group$"0.5"
dfUngra = group$"-0.5"

expGra <- clmm(as.factor(RatingRATING) ~ Trial + Case_contrast + (1+ Case_contrast |Subject)+(1+ Case_contrast |Item), data=dfGra)
summary(expGra) #no effect

expUngra <- clmm(as.factor(RatingRATING) ~ Trial + Case_contrast + (1+ Case_contrast |Subject)+(1+ Case_contrast |Item), data=dfUngra)
summary(expUngra) #sig effect of case, unexpected

## In all the models with random slopes above, the coefficients turn out to be NA.
```