---
title: "Andrea Hofmann: Maulwurf Analysis"
author: "Reinhold Kliegl"
date: "2022-08-24 (last revised: `r format(Sys.time())`)"
output:
  html_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
    fig_width: 8
    fig_height: 6
    fig_caption: true
  pdf_document:
    toc: yes
    toc_depth: 3
    number_sections: yes
    fig_width: 7
    fig_height: 5
    fig_caption: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(arrow)
library(lme4)
library(tidyverse)
library(easystats)
library(summarytools)

# respecting color vision deficiency
cbPalette <- c( "#0072B2", "#D55E00", "#009E73", "#CC79A7",
                "#F0E442", "#56B4E9", "#999999", "#E69F00")
```

# Background

The file was prepared on the basis of the dataframe used for the original LMM analysis.  (see https://osf.io/3u62p/ )

The data are part of the Master thesis that recorded children's (age: 4-8 years) reaction times during a task teaching them syntactic relations. 

# Variables

+ `Subj`: Participant ID (renamed from `ID`; random factor)
+ `Item`: Word ID (random factor)
+ `age`:  4 - 8 years 
+ `Block` (within-Subj/within-Item): 
     + 1st Learning
     + 2nd Learning
     + Disruption
     + Recovery
+ `Target`(renamend fom targetness)
     + Non-target syllable 
     + Target syllable
+ `rt`:  response time 

```{r}
dat <- 
  read_feather("./data/Hofmann_Maulwurf_rt.arrow") |> 
  mutate(a1 = age - 6.210813,
         lrt = log(rt)) |> 
  relocate(a1, .before=Block)
```

+ `age` is centered at mean;  using a round number of years (i.e., 6), rather than the mean yields a more meaningul `GM`.  We keep it for compatibility with original fit.
+ Box-Cox analysis suggested a log transformation of reaction times `rt`

# LMM analysis

## Contrasts and indicator variables

Indicator variables for `Target` and `Block` contrasts
     + `trgt` : Target syllable - Non-target syllable
     + `trng` : 2md learning - 1st learning
     + `drpt` : Disruption - 2nd learning
     + `rcvr` : Recovery - Disruption 

```{r}
contrasts(dat$Target) <- contr.sum(2)
contrasts(dat$Block) <- MASS::contr.sdif(4)
mm <- model.matrix(~ 1 + Target*Block, data=dat)
dat$trgt <- mm[,2]
dat$trng <- mm[,3]
dat$drpt <- mm[,4]
dat$rcvr <- mm[,5]
```

## Original maximal LMM

```
lmax.logRT <-
  lmer(logRT ~ Block * Targetness * c_age + (1 + Block * Targetness | anon_ID) + (1 + c_age| Item), 
       control = lmerControl(calc.derivs = FALSE, optCtrl=list(maxfun=100000), optimizer="bobyqa"), data = SRT.analysis)
```

```{r}
load("./fits/lmax.logRT.RData")
VarCorr(lmax.logRT)
```


## Complex LMM

```{r}
system.time(
m_cpx <- lmer(lrt ~ 1 + trgt * (trng+drpt+rcvr) * a1  + 
                   (1 + trgt * (trng+drpt+rcvr) | Subj) + (1 + a1 | Item),
               data=dat, REML=FALSE, control = lmerControl(calc.derivs = FALSE))
)
VarCorr(m_cpx)
```

```
   user  system elapsed 
 54.186   8.368  56.758 
```

## Parsimonious LMM

```{r}
system.time(
m_prm2 <- lmer(lrt ~ 1 + trgt * (trng+drpt+rcvr) * a1  + 
                    (1 + drpt + rcvr + trng + drpt:trgt + rcvr:trgt | Subj) + 
                    (0 + trgt || Subj) +  (1 + a1 | Item),
               data=dat, REML=FALSE, control = lmerControl(calc.derivs = FALSE))
)
VarCorr(m_prm2)
```

```
   user  system elapsed 
 20.009   2.432  20.612 
```
## Diagnostics

```{r}
compare_models(m_cpx, m_prm2, effects="fixed", style="se_p")

check_model(m_prm2)
```

# Appendix

```{r}
sessionInfo()
```

