---
title: "Maulwurf_Preprocessing_Analysis"
author: "Andrea Hofmann"
date: "02.09.2021 updated `r format(Sys.time(), '%d %B, %Y')"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vroom)
library(lme4)
library(lmerTest)
library(car)
library(ggplot2)
library(ggeffects)
library(dplyr)
library(jtools)
library(sjPlot)
```
  
# Import data sets  

```{r import data}
# All ---------------------------------
# data from all participants that completed at least the SRT task
data <- vroom("./data/maulwurf.csv", col_types = ("fnnfnfffnfffiiffnff"))
# Included ---------------------------------
# subset only with participants included in the analysis
incl <- vroom("./data/maulwurf_incl.csv", col_types = ("fnnfnfffnfffiiffnff"))
# SRT all ---------------------------------
# subset for included participants but only SRT task(correct and incorrect responses)
# here information about item durations in ms included as well
SRT <- vroom("./data/incl_SRT.csv", col_types = ("ffnnfnfnfnffnnffnffnn"))
# SRT correct ---------------------------------
# subset for included participants but only SRT task and only correct responses
# here information about item durations in ms included as well
corrSRT <- vroom("./data/incl_corr_SRT.csv", col_types = ("ffnnfnfnfnffnnffnffnn"))
# SC ---------------------------------
# subset for included participants but only SC task
# two participants less included here bc they left the experiment after the SRT task
sC <- vroom("./data/incl_SC.csv", col_types = ("fnnfnfffnfffiiffnff"))
```
  
# Preprocessing  

## Outlier exclusion  

### First step: exclude based on zeros, RTs within 20 ms, upper and lower cutOff value  

```{r outlier CutOff rt}
# Remove zero values ---------------------------------
# 0 RTs appear bc of a coding error within the experiment
corrSRT$cutOffZero <- (corrSRT$rt == 0)
outlier_zero <- subset(corrSRT, corrSRT$cutOffZero == "TRUE")
percOut_zero <- round(length(outlier_zero$rt)/length(corrSRT$rt)*100,1)

# Remove RTs before stimulus onset ---------------------------------
# 20ms is the pause at the beginning of each stimulus
corrSRT$cutOffStimOnset <- corrSRT$rt > 0 & corrSRT$rt < 21
outlier_StimOnset <- subset(corrSRT, corrSRT$cutOffStimOnset == "TRUE")
percOut_StimOnset <- round(length(outlier_StimOnset$rt)/length(corrSRT$rt)*100,1)

# Remove RTs that fall within motor response planning ---------------------------------
# up to 200ms motor response planning + execution
corrSRT$cutOffRespPlan <- (corrSRT$rt < 221 & corrSRT$rt > 21)
outlier_RespPlan <- subset(corrSRT, corrSRT$cutOffRespPlan == "TRUE")
percOut_RespPlan <- round(length(outlier_RespPlan$rt)/length(corrSRT$rt)*100,1)

# Upper cutOff ---------------------------------
# each upper cutOff is the item duration plus 7 sec
# everything above this value will be excluded from the analysis
corrSRT$Upper <- corrSRT$item_dur + 7000
corrSRT$cutOffUpper <- corrSRT$rt > corrSRT$Upper
outlier_TimeOut <- subset(corrSRT, corrSRT$cutOffUpper == "TRUE")
percOut_TimeOut <- round(length(outlier_TimeOut$rt)/length(corrSRT$rt)*100,1)

# general cutOff column with boolean for cutOff = yes / no
corrSRT$cutOff <- corrSRT$cutOffZero == TRUE | corrSRT$cutOffStimOnset == TRUE | corrSRT$cutOffRespPlan == TRUE | corrSRT$cutOffUpper == TRUE

# Outlier subsets to check amounts ---------------------------------
outlier <- subset(corrSRT, corrSRT$cutOff == "TRUE")
# create subsets of outliers per block to check how many data points were excluded per block
outliers_cutOff1 <- subset(outlier, Block_Nr == "1")
outliers_cutOff2 <- subset(outlier, Block_Nr == "2")
outliers_cutOff3 <- subset(outlier, Block_Nr == "3")
outliers_cutOff4 <- subset(outlier, Block_Nr == "4")
# Total outliers removed
Out_cutOff <- round(length(outlier$rt)/length(corrSRT$rt)*100,1)

# Subset without outliers for use in next step ---------------------------------
outlier_false <- subset(corrSRT, corrSRT$cutOff == "FALSE")

# Clean workspace ---------------------------------
rm(outlier, outlier_RespPlan, outlier_StimOnset, outlier_TimeOut, outlier_zero, outliers_cutOff1, outliers_cutOff2, outliers_cutOff3, outliers_cutOff4, Out_cutOff, percOut_RespPlan, percOut_StimOnset, percOut_TimeOut, percOut_zero)
```  
  
### Exclude outliers as Lammertink et al. did in their script - but 2.5SDs above the mean for each anon_ID per Block and Target  

```{r outlier Lammertink SD, include=FALSE, message=FALSE, warning=FALSE}
# Data frame with means and sds in different aggregations ---------------------------------
# Create dataframe
sd.out <- aggregate(rt ~ anon_ID + Block_Nr + Target, outlier_false, sd)$rt
means.out <- aggregate(rt ~ anon_ID + Block_Nr + Target, outlier_false, base::mean)
# Appropriate naming
means <- means.out$rt
Block_Nr <- means.out$Block_Nr
Target <- means.out$Target
anon_ID <- as.character(means.out$anon_ID)
# Create data frame with aggregated data from above
outliers_SD <- data.frame (
  Block_Nr = Block_Nr,
  anon_ID = anon_ID,
  Target = Target,
  m = means,
  sd = sd.out
)
# Merge data ---------------------------------
outliers.SD <- merge(outlier_false,outliers_SD, by=c("anon_ID", "Block_Nr", "Target"))
# Second step in outlier removal based on SD criterion ---------------------------------
# Remove outliers that are above or below 2.5 SD for each anon_ID in each Block for each Target 
outliers.SD$outlier <- outliers.SD$rt > outliers.SD$m + 2.5*outliers.SD$sd |  outliers.SD$rt < outliers.SD$m - 2.5*outliers.SD$sd
# Outlier subsets to check amounts---------------------------------
outliersSD_true <- subset(outliers.SD, outlier == "TRUE")
# create subsets of outliers per block to check how many data points were excluded per block
outliers_SD1 <- subset(outliersSD_true, Block_Nr == "1")
outliers_SD2 <- subset(outliersSD_true, Block_Nr == "2")
outliers_SD3 <- subset(outliersSD_true, Block_Nr == "3")
outliers_SD4 <- subset(outliersSD_true, Block_Nr == "4")
# Total outliers removed
Out_SD <- round(length(outliersSD_true$rt)/length(outlier_false$rt)*100,1)
# Subset without outliers for use in analysis ---------------------------------
outliersSD_false <- subset(outliers.SD, outlier == "FALSE")
# Final analysis data set for SRT task ---------------------------------
SRT.analysis <- subset(outliersSD_false, Block_Nr == "1" | Block_Nr == "2" | Block_Nr == "3" | Block_Nr == "4")
# Clean workspace ---------------------------------
rm(outlier_false, sd.out, means.out, means, Block_Nr, Target, anon_ID, outliers_SD, outliers.SD, outliersSD_true, outliers_SD1, outliers_SD2, outliers_SD3, outliers_SD4, Out_SD, outliersSD_false)
```  
  
# Online measure - word monitoring SRT task  
  
## Check for data transformation  

```{r transformation checks}
# BoxCox to check for need of transformation ---------------------------------
out <- MASS::boxcox(rt ~ Block_Nr * Target * age + Target * Block_Nr * anon_ID + age * Item, data = SRT.analysis)
range(out$x[out$y > max(out$y)-qchisq(0.95,1)/2])
out$x[which.max(out$y)]
# log transform rts
SRT.analysis$logRT <- log(SRT.analysis$rt)
rm(out)

'hist(SRT.analysis$rt, # histogram
 col="peachpuff", # column color
 border="black",
 prob = TRUE, # show densities instead of frequencies
 xlab = "rt",
 main = "Density")
lines(density(SRT.analysis$rt), # density plot
 lwd = 2, # thickness of line
 col = "chocolate3")

hist(SRT.analysis$logRT, # histogram
 col="peachpuff", # column color
 border="black",
 prob = TRUE, # show densities instead of frequencies
 xlab = "rt",
 main = "Density")
lines(density(SRT.analysis$logRT), # density plot
 lwd = 2, # thickness of line
 col = "chocolate3")'
```  
  
### Analysis  

**Linear mixed model analysis**  
```{r contrast coding and analysis}
# Predictor Block_Nr ---------------------------------
## Repeated contrasts =================================
### Repeated contrasts successively test neighboring factor levels against each other
### Block 2-1 = 2nd Learning - 1st Learning = trainingEffect
### Block 3-2 = Disruption - 2nd learning = disruptionEffect
### Block 4-3 = Recovery - Disruption = recoveryEffect
# characterize
SRT.analysis$Block_Nr <- as.character(SRT.analysis$Block_Nr)
# apprpiate naming
SRT.analysis$Block_Nr[SRT.analysis$Block_Nr %in% "1"] <- "1st Learning"
SRT.analysis$Block_Nr[SRT.analysis$Block_Nr %in% "2"] <- "2nd Learning"
SRT.analysis$Block_Nr[SRT.analysis$Block_Nr %in% "3"] <- "Disruption"
SRT.analysis$Block_Nr[SRT.analysis$Block_Nr %in% "4"] <- "Recovery"
# factorize
SRT.analysis$Block <- as.factor(as.character(SRT.analysis$Block_Nr))
SRT.analysis$Block <- factor(SRT.analysis$Block_Nr)
### Contrast coding #############################
contrasts(SRT.analysis$Block) <- MASS::contr.sdif(4)
# apprpriate naming for effects
colnames(contrasts(SRT.analysis$Block)) <- c("trainingEffect","disruptionEffect", "recoveryEffect")
# this is what is looks like
'             trainingEffect disruptionEffect recoveryEffect
1st_Learning          -0.75             -0.5          -0.25
2nd_Learning           0.25             -0.5          -0.25
Disruption             0.25              0.5          -0.25
Recovery               0.25              0.5           0.75'

# Predictor / covariate Age in years ---------------------------------
## Mean center =================================
### Mean is at 0
SRT.analysis$age <- as.numeric(SRT.analysis$age)
SRT.analysis$c_age<-scale(SRT.analysis$age, center = TRUE, scale = FALSE)

# Predictor Target ---------------------------------
## Sum contrast / 2 =================================
### Estimates difference between target and non-target
### Mean of these contrasts is grand mean
### Target present = target item = + 0.5
### Target absent = non-target item = - 0.5
SRT.analysis$Target <- as.character(SRT.analysis$Target)
SRT.analysis$Target[SRT.analysis$Target %in% "present"] <- "Target syllable"
SRT.analysis$Target[SRT.analysis$Target %in% "absent"] <- "Non-target syllable"
SRT.analysis$Targetness <- factor(SRT.analysis$Target)
contrasts(SRT.analysis$Targetness) <- c(-1/2,1/2)
# apprpriate naming for effect
colnames(contrasts(SRT.analysis$Targetness)) <- c("targetEffect")
# this is what is looks like
'           targetEffect
Non-target         -0.5
Target              0.5'

# Maximal model ---------------------------------
lmax.logRT <-
  lmer(logRT ~ Block * Targetness * c_age + (1 + Block * Targetness | anon_ID) + (1 + c_age| Item), 
       control = lmerControl(calc.derivs = FALSE, optCtrl=list(maxfun=100000), optimizer="bobyqa"), data = SRT.analysis)
```
  
**Check normality assumption (with QQplot) for Log RT model**  

**QQplot logRT model**  

```{r Plot residuals of the model to check model assumptions, echo = FALSE}
# there are still a lot of very long RTs that distort the distribution
qqPlot(residuals(lmax.logRT))
```
  
**Summary of model estimates**  
```{r summary logRT model}
summary(lmax.logRT)
save(lmax.logRT, file="lmax.logRT.Rdata")
# load("lmax.logRT.Rdata")
sigma(lmax.logRT)
sjPlot::tab_model((lmax.logRT), show.se = TRUE, show.stat = TRUE, file="lmaxlogRT.doc")
```  
  
## Plot effects  

```{r plot marginal effects}
# marginal main effects ------------------------------------------
## Block =================================
pred_block <- ggpredict(lmax.logRT, c("Block"))
plot(pred_block, connect.lines = TRUE)+
  labs(
    x = "Block", 
    y = "Predicted mean values of log scaled RTs",
  ) +
  theme(axis.text=element_text(size=11),
        axis.title.x=element_text(size=11),
        axis.title.y=element_text(size=11),
  legend.justification = c("right", "top"),
  legend.title = element_text(size=11, face="bold"), legend.text=element_text(size=12, face="bold")
  )
ggsave("./figures/Block_effect.jpg", width = 18, height = 14, dpi=300, units = "cm")

## Age =================================
pred_age <- ggpredict(lmax.logRT, c("c_age"))
plot(pred_age) + theme_apa() + 
  labs(
    x = "Age (mean centered)", 
    y = "Predicted mean values of log scaled RTs", 
    title = "",
    colour = "Age (mean centered)"
  )
ggsave("Age_Effect.jpg", width = 20, height = 14, dpi=300, units = "cm")

# marginal interactions ------------------------------------------
## age:targetness =================================
pred_age_tar <- ggpredict(lmax.logRT, c("c_age", "Targetness"))
plot(pred_age_tar) +
  labs(
    x = "Age (mean centered)", 
    y = "Predicted mean values of log scaled RTs", 
    title = "",
    colour = "Type"
  )
ggsave("./figures/age_target.png", width = 20, height = 14, units = "cm")

## targetness:age =================================
pred_tar_age <- ggpredict(lmax.logRT, c("Targetness", "c_age"))
plot(pred_tar_age, connect.lines = TRUE) +
  labs(
    x = "Target type", 
    y = "Predicted mean values of log scaled RTs", 
    title = "",
    colour = "Age (mean centered)"
  ) + theme_apa() +
  theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14,face="bold"),
    panel.background = element_rect(fill = "white") # bg of the panel
    , plot.background = element_rect(fill = "white") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank() # get rid of minor grid
    , legend.background = element_rect(fill = "transparent") # get rid of legend bg
    , legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
    , legend.position = c(.59, .99),
  legend.justification = c("right", "top"),
  legend.title = element_text(size=14, face="bold"), legend.text=element_text(size=14, face="bold")
  )
ggsave("./figures/target_age.png", width = 20, height = 14, dpi=300, units = "cm")
```  
  
# Descriptive plots for response time data (SRT task)  
  
## Log RTs by target  

```{r log RT per target}
# Prepare aggregated data ---------------------------------
logRT.summary <- 
  SRT.analysis %>% 
  group_by(Block_Nr, Target) %>%
    summarize(ysd = sd(logRT), ymean = mean(logRT), n=n(), yse = ysd/sqrt(n))

require(ggsignif)
require(ggpubr)
# Plot ---------------------------------
logRT.plot <- ggplot(logRT.summary, aes(x=Block_Nr, y=ymean, colour=Target,linetype=Target, group=Target)) + 
  geom_point(size = 2, position = position_dodge(.4)) +
  geom_errorbar(aes(ymin = ymean-yse, ymax = ymean+yse), position = position_dodge(.4), width=.5) + 
  geom_line(aes(linetype=Target), position=position_dodge(.4)) +
  scale_linetype_manual(values=c("dashed", "solid")) +
  scale_color_grey(end = 0) +
  xlab("") +
  ylab("Mean RT from stimulus onset (log scale)")  + 
  theme_apa() +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12),
        panel.background = element_rect(fill = "white") # bg of the panel
        , plot.background = element_rect(fill = "white") # bg of the plot
        , panel.grid.major = element_blank() # get rid of major grid
        , panel.grid.minor = element_blank() # get rid of minor grid
        , legend.background = element_rect(fill = "transparent") # get rid of legend bg
        , legend.title = element_blank()) + 
  stat_summary(fun=mean, colour="black", geom="text", show.legend = FALSE, 
               vjust=0.5, hjust=-0.35, aes( label=round(..y.., digits=3)), position=position_dodge(.4), size=3) +
  geom_bracket(
    xmin = c(1, 2, 3), xmax = c(2, 3, 4),
    y.position = c(7.78, 7.78, 7.78), label = c("** p < .009", "* p < .05", "n.s."),
    tip.length = 0.03, inherit.aes = FALSE, label.size = 3
  )
ggsave("./figures/Figure_2.jpg", logRT.plot, width = 18, height = 12, dpi = 300, units = "cm")
```  
  
## RTs per age group  

```{r RT plot age}
# Prepare age grouping ---------------------------------
SRT.analysis$age_group <- ifelse(SRT.analysis$age < 4.00, '3-year-old', ifelse(SRT.analysis$age < 5.00, '4-year-olds', ifelse(SRT.analysis$age < 6.00, '5-year-olds',ifelse(SRT.analysis$age < 7.00, '6-year-olds', ifelse(SRT.analysis$age < 8.00, '7-year-olds', '8-year-old')))))
# subset age groups
fourYear <- subset(SRT.analysis, SRT.analysis$age_group =="4-year-olds")
fiveYear <- subset(SRT.analysis, SRT.analysis$age_group =="5-year-olds")
sixYear <- subset(SRT.analysis, SRT.analysis$age_group =="6-year-olds")
sevenYear <- subset(SRT.analysis, SRT.analysis$age_group =="7-year-olds")
eightYear <- subset(SRT.analysis, SRT.analysis$age_group =="8-year-old")
# Plot ---------------------------------
meanAlllogRT.plot <- ggplot(SRT.analysis, aes(x=logRT, y=age_group)) + 
    geom_boxplot() +
  xlab("RT from stimulus onset (log scale)") +
  ylab("")  + theme_bw(base_size = 9,base_family = "Helvetica") +
  theme(axis.text=element_text(size=14, face="bold"),
        axis.title=element_text(size=14,face="bold"),
    panel.background = element_rect(fill = "white") # bg of the panel
    , plot.background = element_rect(fill = "white") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank() # get rid of minor grid
    , legend.background = element_rect(fill = "transparent") # get rid of legend bg
    , legend.box.background = element_rect(fill = "transparent") # get rid of legend panel bg
  ) +
  facet_grid(rows = vars(Block_Nr), scales="free_x") +
    theme(strip.text.x = element_text(size = 14),
          strip.text.y = element_text(size = 14))
ggsave("./figures/boxplotanon_ID.png", width = 18, height = 14, dpi=300, units = "cm")
```  
  
# SC model (accuracy data from stem completion task)  

**Generalized mixed model analysis - logit link function**  

```{r acc analysis, eval=TRUE}
# Appropriate naming for predictor levels ---------------------------------
# characterize
sC$Target <- as.character(sC$Target)
# rename
sC$Target[sC$Target %in% "present"] <- "Target syllable"
sC$Target[sC$Target %in% "absent"] <- "Non-target syllable"
# factorize
sC$Target <- as.factor(as.character(sC$Target))
# characterize
sC$Familiarity <- as.character(sC$Familiarity)
# rename
sC$Familiarity[sC$Familiarity %in% "new"] <- "Novel verb stem"
sC$Familiarity[sC$Familiarity %in% "familiar"] <- "Familiar verb stem"
# factorize
sC$Familiarity <- as.factor(as.character(sC$Familiarity))

# Contrast coding ---------------------------------
## Predictor Familiarity =================================
### Sum / 2 contrasts #############################
### Tests the difference between factor levels
### Mean is at the grand mean
sC$Familiar <- as.factor(sC$Familiarity)
contrasts(sC$Familiar) <- c(1/2,-1/2)
## Predictor / covariate age =================================
### mean center #############################
### overall comparisons for average effects
sC$age <- as.numeric(sC$age)
sC$c_age<-scale(sC$age, center = TRUE, scale = FALSE)
## Predictor Targetness =================================
#### Sum / 2 contrasts #############################
### Tests the difference between factor levels
### Mean is at the grand mean
sC$Targetness <- factor(sC$Target)
contrasts(sC$Targetness) <- c(-1/2,1/2)
# Modelling ---------------------------------
sCAcc <- glmer(Accuracy ~ Targetness * c_age + Familiar * c_age + (1 + Targetness | anon_ID), control=glmerControl(optimizer="bobyqa"), data = sC, family = binomial)
## Model output =================================
summary(sCAcc)
save(sCAcc, file="sCAcc.Rdata")
tab_model(sCAcc, file="sCOdds.doc")
tab_model(sCAcc, transform = NULL, auto.label = FALSE, file="sClogOdds.doc")
sjPlot::tab_model(sCAcc, show.se = TRUE, show.stat = TRUE, file="sCAcc.doc")
```  
  
### Calculate CIs  

```{r compute confidence intervals Accuracy model intercept, warning=FALSE}
CIs.sCAcc <- data.frame()
pvalues.sCAcc <- data.frame()
Intercept.profile <- profile(sCAcc, which = "(Intercept)")
Intercept.CI <- confint(Intercept.profile)
CIs.sCAcc <- rbind(CIs.sCAcc, Intercept.CI)

Targetness1.profile <- profile(sCAcc, which = "Targetness1")
Targetness1.CI <- confint(Targetness1.profile)
CIs.sCAcc <- rbind(CIs.sCAcc, Targetness1.CI)

c_age.profile <- profile(sCAcc, which = "c_age")
c_age.CI <- confint(c_age.profile)
CIs.sCAcc <- rbind(CIs.sCAcc, c_age.CI)

Familiar1.profile <- profile(sCAcc, which = "Familiar1")
Familiar1.CI <- confint(Familiar1.profile)
CIs.sCAcc <- rbind(CIs.sCAcc, Familiar1.CI)

Targetness1c_age.profile <- profile(sCAcc, which = "Targetness1:c_age")
Targetness1c_age.CI <- confint(Targetness1c_age.profile)
CIs.sCAcc <- rbind(CIs.sCAcc, Targetness1c_age.CI)

c_ageFamiliar1.profile <- profile(sCAcc, which = "c_age:Familiar1")
c_ageFamiliar1.CI <- confint(c_ageFamiliar1.profile)
CIs.sCAcc <- rbind(CIs.sCAcc, c_ageFamiliar1.CI)

colnames(CIs.sCAcc) <- c("CIlowLogodds", "CIupLogodds")
CIs.sCAcc$CIlowLogodds <- round(CIs.sCAcc$CIlowLogodds,2)
CIs.sCAcc$CIupLogodds <- round(CIs.sCAcc$CIupLogodds,2)
```

**Odds and probabilities for fixed effects accuracy model**  

```{r Accuracy model compute percentage for glmer intercept (plus range confidence interval)}
## Log-odds (= model estimate)
## odds = exp(log-odds) or odds=P/(1-P)
## probability (P) = odd/(1+odds)

Probability <- function (odds) {
  odds / (1+odds)
}

# Calculate odd-ratio's for each effect

table_acc <- as.data.frame(fixef(sCAcc))
colnames(table_acc) <- "Logodds"
table_acc$Logodds <- round(table_acc$Logodds,2)
table_acc$odds <- round(exp(table_acc$Logodds),2)
table_acc <- cbind(table_acc,CIs.sCAcc)
colnames(table_acc) <- c("Logodds", "odds", "CIlowLogodds", "CIupLogodds")

table_acc$CIlowOdds <- round(exp(table_acc$CIlowLogodds),2)
table_acc$CIupOdds <- round(exp(table_acc$CIupLogodds),2)

print(table_acc)

# Calculate probability for intercept only
Prob_intercept <- round((Probability(table_acc$odds[1])*100),1)
Prob_intercept_CIL <- round((Probability(table_acc$CIlowOdds[1])*100),1)
Prob_intercept_CIH <- round((Probability(table_acc$CIupOdds[1])*100),1)

paste("Probability interecept = ",Prob_intercept,"%",sep="")
paste("Probability interecept 95% CI low = ",Prob_intercept_CIL,"%",sep="")
paste("Probability interecept 95% CI high = ",Prob_intercept_CIH,"%",sep="")
```
  
## Plot effects  

```{r plot sc model predictions}
# Plot all coefficients plus intercept with probabilities ---------------------------------
pred_scAcc <- plot_model(sCAcc, show.intercept = TRUE, scale = TRUE, colors = "bw", show.values = TRUE, show.p=TRUE, transform = "plogis", expand.grid = TRUE) +
  theme_sjplot2() + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=12),
    panel.background = element_rect(fill = "white") # bg of the panel
    , plot.background = element_rect(fill = "white") # bg of the plot
    ) +
  scale_x_discrete(labels=list
                   (
    Targetness1 = "Target \nvs.\n non-target syllable", 
    c_age = "Age (mean centered)",
    Familiar1 = "Familiar \nvs.\n novel verb stem",
    'Targetness1:c_age' = "Target / non-target syllable \n: \nage (mean centered)",
    'c_age:Familiar1' = "Familiar / novel verb stem \n: \nage (mean centered)"
    )
  )
ggsave("./figures/sc_accuracy_model.jpg", pred_scAcc, width = 18, height = 14, dpi = 300, units = "cm")

# Odds ratio ---------------------------------
plot_model(sCAcc, show.values = TRUE, value.offset = .3)
```  
  
# Descriptive plots for accuracy data (stem completion task)  

```{r plot A overall accuracy B familiarity C targetness scores, echo=FALSE, warning=FALSE}
require(gridExtra )
# Figure A: overall response accuracy ---------------------------------
## Aggregate =================================
sC$Accuracy <- as.numeric(as.character(sC$Accuracy))
overall.acc <- aggregate(Accuracy ~ anon_ID, sC, mean)
## Prepare plot =================================
plot.acc <- ggplot(overall.acc, aes(x=0, y=Accuracy)) +
  labs(x= "All responses", y = "Proportion utterances correct") +
  theme_bw(base_size = 12) + xlim(-1/2,1/2)+ ylim(0,1)+ theme(
    panel.background = element_rect(fill = "white") # bg of the panel
    , plot.background = element_rect(fill = "white") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank()# get rid of minor grid
    , axis.text.x = element_blank()
    , axis.ticks.x = element_blank()
    #,axis.title.x = element_blank()
    , legend.background = element_rect(fill = "transparent") # get rid of legend bg
    , legend.box.background = element_rect(fill = "transparent")) # get rid of legend panel bg
## violine plot overall =================================
violin <- plot.acc + geom_violin(trim = FALSE, fill='lightgrey', color = "black", size=0.5) +
  geom_dotplot(binaxis = "y", stackdir = "center", fill="white", dotsize=0.2, alpha = 0.8)  + 
  guides(size= "none") + 
  stat_summary(fun.data = mean_se, geom = "errorbar", size=0.4, width = 0.08) +
  stat_summary(fun = mean,
               geom = "point", size=6, shape=4) + 
  stat_summary(fun.y=mean, colour="black", geom="text", show_guide = FALSE, 
               vjust=0.26, hjust=-0.42, aes( label=round(..y.., digits=2))) +
  annotate("text", x = -0.42, y = 1, label = "A",cex=8)
# Figure B: familiarity ---------------------------------
## Aggregate =================================
overall.acc.gen <- aggregate(Accuracy ~ anon_ID + Familiarity, sC, mean)
overall.acc.gen$Familiarity <- factor(overall.acc.gen$Familiarity, levels=c("familiar", "new"), labels=c("Familiar \nverb stem", "Novel \nverb stem"))
## Prepare plot =================================
plot.acc.fam <- ggplot(overall.acc.gen, aes(x= Familiarity,y=Accuracy)) + 
  labs(x= "", y = "Proportion utterances correct") +
  theme_bw(base_size = 12) + ylim(0,1)+ 
  theme(
    panel.background = element_rect(fill = "white") # bg of the panel
    , plot.background = element_rect(fill = "white") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank()# get rid of minor grid
    , axis.text.x = element_text(size=12)
    , axis.ticks.x = element_blank()
    ,axis.title.x = element_blank()
    , legend.background = element_rect(fill = "transparent") # get rid of legend bg
    , legend.box.background = element_rect(fill = "transparent"),
    axis.title.y=element_blank()
    ) # get rid of legend panel bg
## violine plot familiarity =================================
violin.fam <- plot.acc.fam + 
  geom_violin(trim = FALSE, fill='lightgrey', color = "black", size=0.5) +
  geom_dotplot(binaxis = "y", stackdir = "center", fill="white", dotsize=0.2, alpha = 0.8)  + 
  guides(size= "none") + 
  stat_summary(fun.data = mean_se, geom = "errorbar", size=0.4, width = 0.08) +
  stat_summary(fun = mean,
               geom = "point", size=6, shape=4) +
  stat_summary(fun.y=mean, colour="black", geom="text", show_guide = FALSE, 
               vjust=0.26, hjust=-0.42, aes( label=round(..y.., digits=2))) +
  annotate("text", x = 0.58, y = 1, label = "B",cex=8)
# Figure C: targetness ---------------------------------
## Aggregate =================================
overall.acc.tar <- aggregate(Accuracy ~ anon_ID + Target, sC, mean)
overall.acc.tar$Target <- factor(overall.acc.tar$Target, levels=c("absent", "present"), labels=c("Non-target \nsyllable", "Target \nsyllable"))
## Prepare plot =================================
plot.acc.targetness <- ggplot(overall.acc.tar, aes(x= Target,y=Accuracy)) + 
  labs(x= "Stem completion task", y = "Proportion utterances correct") +
  theme_bw(base_size = 12) + ylim(0,1)+ theme(
    panel.background = element_rect(fill = "white") # bg of the panel
    , plot.background = element_rect(fill = "white") # bg of the plot
    , panel.grid.major = element_blank() # get rid of major grid
    , panel.grid.minor = element_blank()# get rid of minor grid
    , axis.text.x = element_text(size=12)
    #, axis.ticks.y = element_blank()
    , axis.title.x = element_blank()
    , legend.background = element_rect(fill = "transparent") # get rid of legend bg
    , legend.box.background = element_rect(fill = "transparent"),
    axis.title.y=element_blank()
    ) # get rid of legend panel bg
## violine plot targetness =================================
violin.targetness <- plot.acc.targetness + 
  geom_violin(trim = FALSE, fill='lightgrey', color = "black", size=0.5) +
  geom_dotplot(binaxis = "y", stackdir = "center", fill="white", dotsize=0.2, alpha = 0.8)  + 
  guides(size= "none") + 
  stat_summary(fun.data = mean_se, geom = "errorbar", size=0.4, width = 0.08) +
  stat_summary(fun = mean,
               geom = "point", size=6, shape=4) +
  stat_summary(fun.y=mean, colour="black", geom="text", show_guide = FALSE, 
               vjust=0.26, hjust=-0.42, aes( label=round(..y.., digits=2))) +
  annotate("text", x = 0.58, y = 1, label = "C",cex=8)
# Merge all plots and save ---------------------------------
violingrid <- grid.arrange(violin,violin.fam, violin.targetness,nrow=1)
ggsave("./figures/Figure_3.jpg", violingrid, width = 18, height = 14, units = "cm")
```
  
# Learning score  

```{r learning score}
# remove p12 - left the sc task after two items
# remove p10 - left after SRT task
SRT.analysis <- subset(SRT.analysis, SRT.analysis$anon_ID!="p12" & SRT.analysis$anon_ID!="p10")
# Disruption score ---------------------------------
# compute individual disruption scores
# mean of the differences btw 2nd learning block and disruption AND difference btw recovery and disruption
# this can be thought of as the height of the peak (if there is any) for the disruption block
DisruptionScore.df <- aggregate(logRT ~ anon_ID + Block_Nr, SRT.analysis,mean)
DisruptionRegulars.df <- subset(DisruptionScore.df, Block_Nr == "2nd Learning" | Block_Nr == "Recovery")

DisruptionRegulars.df <- aggregate(logRT ~ anon_ID, DisruptionRegulars.df,mean)
colnames(DisruptionRegulars.df) <- c("anon_ID", "logRTregular")

DisruptionRandom.df <- subset(DisruptionScore.df, Block_Nr == "Disruption")
DisruptionRandom.df <- subset(DisruptionRandom.df, select = c("anon_ID", "logRT"))
colnames(DisruptionRandom.df) <- c("anon_ID", "logRTrandom")

DisruptionScore.df <- merge(DisruptionRandom.df,DisruptionRegulars.df, by = c("anon_ID"))
DisruptionScore.df$DisruptionScore <- DisruptionScore.df$logRTrandom - DisruptionScore.df$logRTregular
DisruptionScore.df <- subset(DisruptionScore.df, select = c("anon_ID", "DisruptionScore"))

# Accuracy score ---------------------------------
# compute individual accuracy scores for the stem completion task
SCScore.df <- aggregate(Accuracy ~ anon_ID, sC, mean)
colnames(SCScore.df) <- c("anon_ID", "SC_Score")

# Learning score ---------------------------------
LearningScores.df <- merge(DisruptionScore.df, SCScore.df, by= c("anon_ID"))
LearningScores.df$DisruptionScore <- round(LearningScores.df$DisruptionScore,4)
LearningScores.df$SC_Score <- round(LearningScores.df$SC_Score,4)

# Correlation plot ---------------------------------
Correlate.plot <- ggplot(LearningScores.df, aes(x=DisruptionScore, y=SC_Score)) +
    geom_point(shape=3, size=2, color="darkred") +    # Use hollow circles
    geom_smooth(method=lm, se = FALSE) + # Add linear regression line # (by default includes 95% CI)
    xlab("Disruption Peak") +
  ylab("Response accuracy in stem completion task")  + theme_bw(base_size = 18, base_family = "Helvetica") + theme(axis.text.x = element_text(size=18),axis.text.y = element_text(size=18))
print(Correlate.plot)
# Pearson correlation ---------------------------------
cor.test(LearningScores.df$DisruptionScore, LearningScores.df$SC_Score)
```

**Comparison of SRT disruption peak and SC accuracy score by means of a model comparison**  
```{r correlation SRT and SC learning scores via model comparison, echo =TRUE}
# Merge learning scores with SRT.analysis ---------------------------------
SRT.analysis <- merge(SRT.analysis, SCScore.df, by = c("anon_ID"))
# Scale learning score ---------------------------------
## Scale SC score as it will be added as a continuous predictor to the RT model
SRT.analysis$SC_Score_scaled <- scale(SRT.analysis$SC_Score)
# Run models ---------------------------------
## Run RT model again with REML = false (otherwise models can't be compared)
SRT.analysisLmm <- lmer(logRT ~ Block * Targetness * c_age + (1 + Block * Targetness | anon_ID) + (1 + c_age | Item), control = lmerControl(optCtrl=list(maxfun=100000),calc.derivs = FALSE, optimizer="bobyqa"), REML=FALSE, data = SRT.analysis)
## Run RT model plus learning score as predictor
SRT.analysisLmmPlus <- lmer(logRT ~ Block * Targetness * c_age + SC_Score_scaled + (1 + Block * Targetness | anon_ID) + (1 + c_age | Item), control = lmerControl(optCtrl=list(maxfun=100000),calc.derivs = FALSE, optimizer="bobyqa"), REML=FALSE, data = SRT.analysis)
# Anova model comparison ---------------------------------
anova(SRT.analysisLmm, SRT.analysisLmmPlus)
```

  